\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{url}
\usepackage{booktabs}
\usepackage{hyperref}

\lstset{
  basicstyle=\ttfamily\footnotesize,
  breaklines=true,
  breakatwhitespace=true,
  columns=flexible,
  keepspaces=true,
  frame=single,
  rulecolor=\color{black!30},
  backgroundcolor=\color{black!4},
  xleftmargin=6pt,
  xrightmargin=6pt,
  aboveskip=4pt,
  belowskip=4pt,
}

% Float environment for code blocks so captions render via LaTeX's
% normal float-caption mechanism (NOT lstlisting's, which clips).
\usepackage{float}
\floatstyle{plain}
\newfloat{codelisting}{tbp}{lop}
\floatname{codelisting}{Listing}

\begin{document}

\title{Two-Feature Detection of WireGuard-in-udp2raw:\\
A Lab Replication and Probe-and-Classify Methodology}

\author{\IEEEauthorblockN{Steven Schiavone}
\IEEEauthorblockA{\textit{Seidenberg School of CSIS} \\
\textit{Pace University}\\
New York, NY \\
sschiavone@pace.edu}
\and
\IEEEauthorblockN{Nicholas Johannsen}
\IEEEauthorblockA{\textit{Seidenberg School of CSIS} \\
\textit{Pace University}\\
New York, NY \\
nj06718n@pace.edu}
}

\maketitle

\begin{abstract}
WireGuard's wire format is so regular that a 25-line classifier identifies it
with negligible false-positive rate. The community response has been
external obfuscation, most prominently \textit{udp2raw}, a single-binary
tunnel that wraps UDP datagrams in crafted faketcp packets and is widely
deployed by users behind censorious networks. To our knowledge, no
peer-reviewed analysis of udp2raw has been published. We present two
elementary attacks on a WireGuard-over-udp2raw deployment and validate them
in a clean lab. (1)~An \textbf{active prober} of five stateless TCP probes
distinguishes a udp2raw faketcp listener from a real Linux kernel TCP
listener in 8.5~seconds with deterministic outcome across five repeated
trials, exploiting that udp2raw's userspace state machine silently drops
non-handshake-shaped packets that a real kernel would answer.
(2)~A \textbf{passive flow-feature classifier} using just two features ---
the fraction of MTU-class ``bulk'' packets and the fraction of 60-byte
ACK-only segments --- separates 24~WireGuard flows (both bare and
udp2raw-wrapped) from 1058~non-VPN flows in the ISCXVPN2016 corpus at 99.9\%
accuracy and ROC AUC of $1.000$. Both features are mechanical consequences
of how the protocols are designed, not statistical regularities that would
be expected to drift across networks. The work positions itself as a small
contribution alongside the recent Xue~\textit{et~al.} line of work on
encapsulated-tunnel fingerprinting, addressing a gap that line has not
covered: a protocol-specific analysis of the udp2raw shim, the most-
downloaded ($\approx$259k binary downloads) generic UDP-to-faketcp
obfuscation tool used in front of WireGuard.
\end{abstract}

\begin{IEEEkeywords}
WireGuard, VPN, deep packet inspection, udp2raw, traffic obfuscation,
flow fingerprinting, active probing
\end{IEEEkeywords}

\section{Introduction}

For VPN users in network environments that perform deep packet inspection,
two distinct properties are at stake: confidentiality of the traffic
carried, and unobservability of the VPN's own existence. The censorship-
circumvention literature, beginning at least with the Tor pluggable
transports work and continuing through Frolov and Wustrow's analysis of
TLS-based circumvention~\cite{frolov2020probe}, treats these as separate
problems requiring separate mechanisms. WireGuard~\cite{donenfeld2017},
designed by J.\ A.\ Donenfeld and shipped as the default VPN in many
recent distributions, addresses confidentiality with a small modern
cryptographic stack and explicitly leaves unobservability out of
scope~\cite{wgknown}. The on-the-wire framing reflects that design
choice: four message types with a constant four-byte header and three
fixed handshake lengths, sufficient for a complete byte-and-length
classifier in roughly twenty lines of Python (we provide one in
\S\ref{sec:trivial-passive}).

The user-facing response, particularly in restrictive networks, has been an
ecosystem of tools that wrap or modify WireGuard's transport. Two
mechanism categories exist, and they are not comparable.
The first category modifies WireGuard itself or replaces parts of its
stack: AmneziaWG~\cite{amneziawg}, swgp-go, and the
\texttt{amneziawg-linux-kernel-module}. The second category is generic
single-binary UDP-to-something shims that are protocol-agnostic and
deployable in front of any UDP service: udp2raw~\cite{udp2raw}, wstunnel,
and Cloak. We focus on the second category. Within it, udp2raw is the
most-downloaded tool whose explicit purpose is wrapping UDP in a
TCP-shaped tunnel: GitHub release statistics~\cite{ghapi_udp2raw}
record approximately $259{,}000$ binary downloads across 33 releases of
the primary repository as of May~2026, with a further $41{,}000$ downloads
of the multiplatform fork. udp2raw is also a documented target of
active blocking by the GFW~\cite{wu2023gfw} and Iranian
DPI~\cite{tegant_amnezia_2024}. Yet, unlike OpenVPN~\cite{xue2022openvpn},
Shadowsocks~\cite{alice2020shadowsocks}, or Tor pluggable transports,
udp2raw has --- to the best of our knowledge, and we welcome correction
--- received no peer-reviewed security analysis. Public mentions are
limited to the WireGuard project's ``Known Limitations''
page~\cite{wgknown} and operational reports from the
GFW Report blog and the Amnezia issue tracker.

This paper closes that gap with two elementary, easy-to-reproduce attacks
and a clean lab evaluation:

\begin{itemize}
  \item \textbf{Active probing of udp2raw faketcp} (\S\ref{sec:c1}). Five
        stateless TCP probes sent at one-second intervals discriminate
        between a real Linux TCP listener and a udp2raw faketcp listener
        with deterministic outcome and in approximately 8.5 seconds. The
        signal is structural: udp2raw's userspace state machine accepts
        and answers a SYN, but silently drops three of four follow-up
        probes that a real kernel always answers.
  \item \textbf{Passive two-feature classification} (\S\ref{sec:c2}).
        WireGuard's MTU-padding habit produces a strikingly high fraction
        of bulk-sized packets (mean 0.55, ${\sigma}$=0.18 across our
        captures) compared to a non-VPN background corpus (mean 0.02,
        ${\sigma}$=0.07). udp2raw's faketcp wrapper additionally produces a
        unique 1:1 data-to-ACK packet ratio: every WireGuard datagram is
        framed as a faketcp data segment that elicits a 60-byte
        kernel-style ACK. Combining these two features yields a
        Random-Forest classifier with 99.9\% accuracy, precision 1.000 and
        ROC AUC 1.000 on a corpus of 24 WireGuard flows and 1058 non-VPN
        flows from ISCXVPN2016~\cite{draper2016iscx}.
\end{itemize}

We frame the contribution narrowly. Recent work by
Xue~\textit{et~al.}~\cite{xue2024obfuscated, xue2025crosslayer} establishes
\emph{protocol-agnostic} detectors for obfuscated proxy tunnels using
encapsulated-TLS-handshake fingerprints (USENIX Security~'24) and
cross-layer round-trip-time anomalies (NDSS~'25). Our results are not
intended to compete with that line. WireGuard's outer encryption nullifies
the encapsulated-TLS signal entirely (no nested TLS handshake exists in a
WireGuard flow), and our captures use UDP-only, short-flow workloads that
attenuate the cross-layer-RTT signal. Protocol-specific structural features
remain the available approach in this regime, and that is what we provide.

The paper proceeds as follows. \S\ref{sec:background} reviews the WireGuard
wire format and udp2raw operation. \S\ref{sec:lab} describes a
fully-reproducible lab. \S\ref{sec:trivial-passive} demonstrates the
trivial passive WireGuard fingerprint that motivates the obfuscation
problem in the first place, and shows udp2raw defeats it. \S\ref{sec:c1}
and \S\ref{sec:c2} present the two new contributions. \S\ref{sec:related}
positions the work against prior literature; \S\ref{sec:limits} is honest
about scope; \S\ref{sec:conclusion} concludes.

\section{Background}
\label{sec:background}

\subsection{The WireGuard wire format}

The WireGuard whitepaper~\cite{donenfeld2017} specifies four message types,
each carried in a single UDP datagram. Every message begins with a one-byte
type identifier and three reserved zero bytes:

\begin{lstlisting}
+---------+-------------------+
| type(1) | reserved zero (3) | ...
+---------+-------------------+
\end{lstlisting}

The handshake messages are also \textit{fixed length}, summarized in
Table~\ref{tab:wgtypes}.

\begin{table}[t]
\centering
\caption{WireGuard UDP payload signature.}
\label{tab:wgtypes}
\begin{tabular}{@{}lll@{}}
\toprule
Type byte & Message & UDP payload length \\
\midrule
0x01 & Handshake Initiation & 148 bytes \\
0x02 & Handshake Response   &  92 bytes \\
0x03 & Cookie Reply         &  64 bytes \\
0x04 & Transport Data       & variable, $\geq 32$ bytes \\
\bottomrule
\end{tabular}
\end{table}

Three concatenated checks --- ``UDP datagram, first byte is one of
$\{1,2,3\}$, next three bytes are zero, length is exactly 148/92/64''---
yield a handshake fingerprint with negligible false-positive risk in normal
traffic. We make explicit in \S\ref{sec:trivial-passive} that this is not
a discovery; it is a re-derivation of a property that is documented in
Donenfeld's own talks and the WireGuard project's mailing list, and that
is reportedly used in production by the GFW~\cite{wu2023gfw}. The
re-derivation matters here only as the motivation for the obfuscation
deployments we then probe.

\subsection{udp2raw and the faketcp transport}

udp2raw~\cite{udp2raw} takes UDP datagrams on one side and re-emits them as
crafted raw packets that imitate other protocols (TCP, ICMP, or fake-UDP).
In its \texttt{faketcp} mode --- the mode we examine, and the strongest of
its disguises --- udp2raw constructs each direction of the tunnel as what
appears to be a real TCP connection. A SYN/ACK three-way handshake is
performed at the start; subsequent traffic is exchanged with plausible
sequence numbers, ACK behaviour, and TCP options.

The \emph{implementation} of this faketcp tunnel is, however, entirely in
userspace. udp2raw uses raw sockets and an AF\_PACKET listener to read
inbound packets out from under the kernel TCP stack. It must do so because
the kernel never opens a real socket on the listening port; if it did, the
kernel would answer SYNs with its own SYN-ACK and emit RST on stray
segments, defeating the disguise. The cost is that udp2raw's state machine
sees \emph{every} packet to its port and must classify each one. As we
show in \S\ref{sec:c1}, it does so non-conformantly with respect to RFC
793/9293: any packet that does not fit udp2raw's small set of expected
shapes is silently dropped, while a real Linux kernel would have answered.

\section{Lab environment}
\label{sec:lab}

The experimental setup, illustrated in Fig.~\ref{fig:topo}, consists of
three nodes:

\begin{enumerate}
  \item A WireGuard server on an OCI VM.\@ Oracle Linux 9.7 on aarch64
        (VM.Standard.A1.Flex), public IP \texttt{X.X.X.X} (redacted in
        public artifacts). WireGuard listens on UDP/51820. For the
        obfuscated path, udp2raw v20230206.0 listens on TCP/4096 and
        forwards to localhost UDP/51820.
  \item A Kali Linux 2025.3 client running in VirtualBox on a Windows 11
        host. Network adapter is NAT (10.0.2.15) since Wi-Fi-to-VM
        bridging is unreliable on common Wi-Fi NIC drivers.
  \item A capture/analysis tool running on the Kali NAT interface, taking
        the role of a passive on-path observer (\S\ref{sec:c2}) or an
        active prober (\S\ref{sec:c1}).
\end{enumerate}

\begin{figure}[t]
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}rl@{}}
\multicolumn{2}{@{}l@{}}{\textbf{(a) Direct path} (passively detectable)}\\
\textit{Kali wg client} & $\xrightarrow{\;\text{UDP / 51820}\;}$ \textit{OCI wg server}\\[6pt]
\multicolumn{2}{@{}l@{}}{\textbf{(b) Obfuscated path} (this paper's target)}\\
\textit{Kali wg client} & $\xrightarrow{\;\text{lo / 51821}\;}$ \textit{Kali udp2raw client}\\
& $\xrightarrow{\;\text{TCP / 4096 (faketcp)}\;}$ \textit{OCI udp2raw server}\\
& $\xrightarrow{\;\text{lo / 51820}\;}$ \textit{OCI wg server}\\
\end{tabular}
\caption{Two end-to-end paths under test. Same WireGuard keys, same
inner addressing (10.66.66.0/24); only the outer transport between
client and server differs.}
\label{fig:topo}
\end{figure}

The client maintains two WireGuard configurations
(\texttt{wg-direct}, \texttt{wg-obfuscated}) using identical keys; only
the \texttt{Endpoint=} differs. This isolates the obfuscation as the sole
independent variable for direct comparison.

\subsection{Path MTU}

Stacking WireGuard inside udp2raw inside Ethernet introduces a non-trivial
MTU constraint. WireGuard adds 80 bytes of overhead; udp2raw faketcp adds
roughly 60 more. With the default WireGuard MTU of 1420 bytes, each WG
transport packet exceeds udp2raw's effective payload budget; udp2raw warns
\texttt{huge packet, data len=1472 (>=1375)} and silently drops the offender,
which is then visible to WireGuard as packet loss. We set MTU=1200 on the
\texttt{wg-obfuscated} interface, after which 30-packet ICMP bursts to the
inner peer succeed at 0\% loss with $\sim$14\,ms RTT. The MTU choice is the
only piece of additional engineering required to deploy the obfuscation; it
is not a contribution but is documented for reproducibility.

All scripts and configurations are bundled in the project repository.

\section{The trivial passive fingerprint (motivation only)}
\label{sec:trivial-passive}

We re-derived the well-known WireGuard byte-and-length fingerprint in a
$\approx$25-line Python predicate (Listing~\ref{lst:classify}) and a
companion Scapy sniffer. This is not a contribution; it is the baseline
that justifies the obfuscation problem.

\begin{codelisting}[t]
\caption{The complete WireGuard fingerprint predicate.}
\label{lst:classify}
\begin{lstlisting}
WG_TYPES = {
    1: ("HANDSHAKE_INIT",     148),
    2: ("HANDSHAKE_RESPONSE",  92),
    3: ("COOKIE_REPLY",        64),
    4: ("TRANSPORT_DATA",    None),
}

def classify(payload):
    if len(payload) < 4:
        return None
    msg_type = payload[0]
    if msg_type not in WG_TYPES \
       or payload[1:4] != b"\x00\x00\x00":
        return None
    label, expected_len = WG_TYPES[msg_type]
    if expected_len is not None:
        if len(payload) != expected_len:
            return None
        return label, msg_type, True
    if len(payload) < 24:
        return None
    return label, msg_type, False
\end{lstlisting}
\end{codelisting}

In a $25$-second capture of a busy \texttt{wg-direct} flow, this predicate
identified all 96~UDP packets in the flow as WireGuard, including the
148-byte handshake initiation and 92-byte handshake response. We sanity-
checked the predicate with eight synthetic test cases (four valid
messages, four near-misses); all eight pass.

The same predicate run against a $25$-second capture of the
\texttt{wg-obfuscated} path produces \texttt{No WireGuard traffic detected}.
The byte-level fingerprint is fully eliminated by the udp2raw wrapper.
The remainder of the paper concerns whether the wrapped tunnel can be
detected by other means, and shows that it can.

\section{Contribution 1: Active probing of udp2raw faketcp}
\label{sec:c1}

\subsection{Threat model and probe design}

We assume an adversary who can issue TCP packets to a target IP and port
and observe replies. This is the same threat model as the Xue~\textit{et~al.}
2022 active-probe stage~\cite{xue2022openvpn} and the
Alice~\textit{et~al.}~\cite{alice2020shadowsocks} active-probe taxonomy for
Shadowsocks. The adversary does not need any traffic from the target user.

We designed five stateless probes, each sending at most one packet and
waiting up to three seconds for a reply. The five probes are:

\begin{enumerate}
  \item[P1] \textbf{Baseline SYN}: a normal SYN with random sequence
        number. Both real TCP and udp2raw answer with SYN-ACK. The probe
        confirms the target is alive and serves as a control.
  \item[P2] \textbf{SYN with payload}: a SYN carrying 32 bytes of payload.
        RFC~9293 forbids application data on a SYN unless TFO is in use; a
        real Linux kernel ignores the payload and answers SYN-ACK normally,
        and the listening service may then echo a banner. udp2raw's parser
        rejects packets that don't fit its expected handshake layout
        and does not answer.
  \item[P3] \textbf{Bogus-ACK}: an ACK packet to the same port carrying
        a wildly wrong sequence/acknowledgment number, with no prior
        connection. RFC~9293 specifies a real TCP stack must answer with
        RST. udp2raw consumes the packet via AF\_PACKET, fails to recognize
        it, and discards it; on some setups the kernel later emits its own
        RST after a delay, but this also occurred on udp2raw in our
        captures (RST in 55\,ms vs.\ 39\,ms for real TCP), so we do not
        rely on the timing of this probe.
  \item[P4] \textbf{Window-scale anomaly}: a SYN with TCP option
        \texttt{WS=14}. Real TCP servers either echo or omit a
        window-scale option in their SYN-ACK; udp2raw's parser rejects
        the option layout and does not answer.
  \item[P5] \textbf{FIN to unopened port}: a FIN packet to a port with no
        established connection. Real TCP stacks answer RST or send no
        reply per RFC; udp2raw silently drops.
\end{enumerate}

The discriminator is the count of \emph{silent} responses among
\{P2, P4, P5\}. Real TCP servers in our experiments answer all three;
udp2raw answers none.

\subsection{Implementation}

We implemented the prober in $\sim$140 lines of Python using
Scapy~\cite{scapy} (Listing~\ref{lst:probe}, abbreviated). It runs in a
single process, sends the five probes sequentially, classifies the
result, and emits a verdict label along with timing detail to a
timestamped evidence file. Evidence files are reproduced verbatim in the
project repository.

\begin{codelisting}[t]
\caption{Active prober verdict logic (abridged; full source in
\texttt{novel/active\_probe.py}).}
\label{lst:probe}
\begin{lstlisting}
def classify(results):
    p = {r["probe"]: r for r in results}
    syn   = p["syn"]["outcome"]
    if syn == "no_response":
        return "UNREACHABLE_OR_FILTERED", ...
    if syn == "rst":
        return "PORT_CLOSED", ...
    silent = sum(1 for x in (
        p["syn_with_payload"]["outcome"],
        p["wscale"]["outcome"],
        p["fin_to_unopened"]["outcome"]
    ) if x == "no_response")
    if syn == "syn_ack" and silent >= 2:
        return "UDP2RAW_FAKETCP_SUSPECTED", \
               f"{silent}/3 follow-up probes silent"
    if syn == "syn_ack" and silent == 0:
        return "REAL_TCP", ...
    return "INCONCLUSIVE", ...
\end{lstlisting}
\end{codelisting}

\subsection{Results}

We ran the prober against three targets on the same OCI host:
TCP/4096 (the udp2raw listener), TCP/22 (the system SSH server, our
real-TCP control), and TCP/9999 (a closed port, our negative control).
Table~\ref{tab:probe} summarizes one run; the udp2raw verdict was
identical across five repeated trials (5/5 \texttt{UDP2RAW\_FAKETCP\_SUSPECTED}).
We did not encounter inconclusive verdicts.

\begin{table*}[t]
\centering
\caption{Active prober per-probe outcomes (one representative run; udp2raw
verdict held across 5/5 repeated trials).}
\label{tab:probe}
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{@{}llll@{}}
\toprule
Probe & TCP/22 (real SSH) & TCP/4096 (udp2raw) & TCP/9999 (closed) \\
\midrule
P1 \textemdash{} SYN                 & SYN-ACK in 181\,ms & SYN-ACK in 163\,ms & no response \\
P2 \textemdash{} SYN + 32\,B payload & SYN-ACK + 2-byte echo & \emph{silent} & no response \\
P3 \textemdash{} bogus-seq ACK       & RST in 81\,ms      & RST in 91\,ms      & no RST \\
P4 \textemdash{} SYN with WS=14      & SYN-ACK            & \emph{silent}      & no response \\
P5 \textemdash{} FIN to unopened     & RST                & \emph{silent}      & no response \\
\midrule
Total wall-clock                     & 0.5\,s             & 8.5\,s             & 15.5\,s \\
Verdict                              & \texttt{REAL\_TCP} & \texttt{UDP2RAW\_FAKETCP\_SUSPECTED} & \texttt{UNREACHABLE\_OR\_FILTERED} \\
\bottomrule
\end{tabular}
\end{table*}

The signal is structural and easy to interpret: the SSH server answers all
five probes promptly, udp2raw answers only P1 and P3, and the closed port
answers nothing within the timeouts. Because the discrimination requires
only counting silent probes among a known set of three, the verdict is
robust to minor timing variation and need not be retuned per-target.

\subsection{Implications and limitations of C1}

The active-probe finding has practical weight: an adversary need not
perform any large-scale traffic capture to identify suspected udp2raw
listeners, only a handful of stateless TCP packets per (IP, port) pair.
The cost of a 5-probe sweep is roughly 1.7\,KB of network traffic and
under 9~seconds wall-clock per target.

We make no claim of novelty for the \emph{technique} of probing for
non-conformant TCP behaviour; this is a well-established class of method
(Xue~\cite{xue2022openvpn}, Frolov~\cite{frolov2020probe},
Alice~\cite{alice2020shadowsocks}). The novelty is the specific
characterisation of udp2raw's response pattern. Two limitations are worth
noting. First, our characterisation is of udp2raw v20230206.0; future
versions may add more conformant TCP-state-machine emulation. We note
that upstream cadence is slow --- the most recent functional release is
\texttt{20230206.0} from February~2023, the most recent functional commit
to the default branch is from June~2024, and as of May~2026 none of the
eight open pull requests touch the faketcp parser logic --- but a
sufficiently motivated maintainer could close the gap. Second, an adversary intent on
producing few false positives must distinguish udp2raw from \emph{other}
non-conformant TCP services, including stateful firewalls in their drop-
silently mode and various honeypots. We have not attempted that
discrimination here.

\section{Contribution 2: Two-feature passive classification}
\label{sec:c2}

\subsection{Hypothesis}

WireGuard's transport packets are AEAD-encrypted with a fixed-length tag
and are typically issued at the wg interface MTU. udp2raw is a stateless
framer: it reads each WG datagram, prepends faketcp framing, and emits a
single TCP segment. It does not shape, batch, or coalesce.

We hypothesised, prior to looking at any aggregate statistics, that this
combination would produce two interpretable structural features:

\begin{itemize}
  \item \textbf{High bulk-packet fraction.} A WireGuard tunnel carrying
        any non-trivial traffic produces many MTU-class packets, because
        WireGuard pads every packet to the configured MTU. Even modest
        web-page loads through the tunnel push large fractions of bulk
        packets in our captures.
  \item \textbf{Anomalous 1:1 ACK ratio under udp2raw.} Because udp2raw
        treats each WG datagram as a faketcp ``data segment,'' the
        receiver's kernel observes one new TCP segment per datagram and
        emits a 60-byte ACK in response. Real bulk TCP, by contrast,
        uses delayed ACK and produces roughly one ACK per two data
        segments. The udp2raw flow's ACK fraction therefore approaches
        0.5 instead of 0.33.
\end{itemize}

We deliberately did not include the protocol's well-known 25-second
\texttt{PersistentKeepalive} cadence as a feature. Initial captures
showed that on a busy tunnel WireGuard does not emit keepalives at all
(traffic is generated by the link itself), so the cadence is unreliable as
a real-world feature.

\subsection{Feature definitions}

For each bidirectional flow keyed by 5-tuple, we computed:

\begin{itemize}
  \item \texttt{bulk\_fraction}: fraction of packets in the flow whose
        IP payload length is $\geq$1200 bytes.
  \item \texttt{ack60\_fraction}: fraction of packets whose IP payload
        length is exactly 60 bytes.
\end{itemize}

We required each flow to span at least 5\,seconds and contain at least
10~packets to be eligible for inclusion. The full feature extractor adds
six additional features (entropy of packet sizes, top-1 and top-3 size
mode fraction, packet rate, ACK-to-data ratio, unique-size count) to
support a comparison between the two-feature minimal model and an
8-feature alternative; we report results for both.

\subsection{Datasets}

\textbf{Positive class (WireGuard, $n=24$).} We collected six independent
90-second captures of \texttt{wg-direct} traffic and six of
\texttt{wg-obfuscated}. The obfuscated captures resolved into 18 distinct
flows because udp2raw's faketcp connection is renegotiated periodically
(roughly every 17~seconds in our network), each renegotiation appearing as
a new TCP 5-tuple. Each \texttt{wg-direct} capture produced exactly one
flow.

\textbf{Negative class (non-VPN, $n=1058$).} We extracted features from
the smallest 23 pcaps in NonVPN-PCAPs-01 of
ISCXVPN2016~\cite{draper2016iscx}, yielding 1058 flows that pass the
duration and packet-count filter. ISCXVPN2016 was selected because (a) it
is canonical in the VPN/non-VPN traffic-classification literature and (b)
it contains a full mix of interactive, file-transfer, streaming, chat,
email, and voice flows in non-VPN form, providing a realistic adversary's
target distribution.

We did not include the dataset's VPN partition, since our positive class is
specifically WireGuard (and OpenVPN/PPTP samples in ISCXVPN2016 would
provide the wrong negative).

\subsection{Results}

The class means of the two headline features (Table~\ref{tab:means}) show
a separation of approximately 29$\times$ in \texttt{bulk\_fraction} and
4$\times$ in \texttt{ack60\_fraction}.

\begin{table}[t]
\centering
\caption{Per-feature class statistics (mean$\pm$std). $n=24$ WG, $n=1058$ background.}
\label{tab:means}
\begin{tabular}{@{}lcc@{}}
\toprule
Feature & WireGuard & Non-VPN background \\
\midrule
\texttt{bulk\_fraction}  & $0.553\pm0.186$ & $0.019\pm0.066$ \\
\texttt{ack60\_fraction} & $0.295\pm0.187$ & $0.069\pm0.169$ \\
\midrule
\texttt{len\_entropy}    & $1.935\pm0.820$ & $1.233\pm1.422$ \\
\texttt{dominant\_size\_fraction} & $0.496\pm0.192$ & $0.710\pm0.304$ \\
\texttt{top3\_size\_fraction}     & $0.877\pm0.150$ & $0.871\pm0.186$ \\
\texttt{ack\_to\_data\_ratio}     & $0.508\pm0.362$ & $0.413\pm2.900$ \\
\texttt{n\_unique\_sizes}         & $13.2\pm 8.5$  & $11.2\pm39.8$ \\
\texttt{rate\_pps}                & $3.23\pm0.90$  & $3.37\pm30.27$ \\
\bottomrule
\end{tabular}
\end{table}

We trained a Random Forest (200~trees, balanced class weight) and a
logistic-regression classifier on (i) the two-feature minimal set and
(ii) the eight-feature full set, with five-fold stratified
cross-validation. Performance is summarised in Table~\ref{tab:perf}.

\begin{table}[t]
\centering
\caption{Cross-validated performance on the WireGuard-vs-non-VPN task.}
\label{tab:perf}
\begin{tabular}{@{}lccccc@{}}
\toprule
Model & Acc.\ & Prec.\ & Rec.\ & F1 & AUC \\
\midrule
RF, 2 features  & 0.9991 & 1.000 & 0.958 & 0.979 & 1.000 \\
LR, 2 features  & 0.9686 & 0.414 & 1.000 & 0.585 & 0.999 \\
RF, 8 features  & 0.9982 & 1.000 & 0.917 & 0.957 & 1.000 \\
LR, 8 features  & 0.9787 & 0.511 & 1.000 & 0.676 & 0.999 \\
\bottomrule
\end{tabular}
\end{table}

The two-feature Random Forest is the headline result: 99.9\% accuracy with
\emph{zero false positives} on 1058 background flows, and 23/24 WG flows
correctly classified at AUC=1.000. Adding the six additional features
slightly degrades RF performance (recall drops from 0.958 to 0.917),
suggesting the additional features add noise without separating signal.
Logistic regression catches every WG flow but pays for it with
$\sim$3\% false-positive rate, consistent with a linear boundary on the
two-feature space being not quite tight enough to exclude the most
keepalive-like background flows.

\subsection{Why the two-feature model wins}

We had initially also expected high \emph{dominant-size-fraction} (the
fraction of packets at the single most common size) to be a strong
WireGuard fingerprint. The data inverted that expectation:
\texttt{dominant\_size\_fraction} for WireGuard is 0.496, but for the
ISCXVPN2016 background it is 0.710 --- the opposite direction. The reason
is that the non-VPN corpus contains many degenerate flows --- TCP
keepalives, DHCP, NetBIOS broadcasts, IPv6 link-local discovery --- whose
packet sizes are essentially constant. WireGuard transport, while heavily
mode-concentrated, mixes 1386-byte and shorter packets and ends up
\emph{less} mode-concentrated than these idle flows. The same applies to
\texttt{top3\_size\_fraction} (no separation) and \texttt{len\_entropy}
(separation in the opposite direction from naive intuition).

This finding is in itself a methodological point: features that look
discriminating against a single-flow self-test can be inverted by the
distribution of degenerate flows in real-world traffic. In our case, the
two features that survived this test are exactly the two that are
mechanically grounded in WireGuard's MTU-padding habit and udp2raw's
1:1 faketcp framing. They are unlikely to drift across networks because
they are consequences of the protocols' implementations, not of the
specific traffic carried.

\subsection{Limitations and ablation discussion of C2}

Our positive class is small ($n=24$) by ML standards, but the variance of
the two key features across captures is unusually low: the six trial
captures of \texttt{wg-direct} produced \texttt{dominant\_size\_fraction}
values of $\{0.805, 0.805, 0.804, 0.789, 0.805, 0.799\}$ and
\texttt{bulk\_fraction} values of $\{0.821, 0.821, 0.820, 0.808, 0.821,
0.815\}$, with ranges of $0.016$ and $0.013$ respectively. This
reproducibility, together with AUC=1.000, suggests a deterministic
structural fingerprint rather than a small-sample fluke.

A natural \emph{defence} against the C2 classifier is to disable bulk
padding: reduce the WireGuard MTU so that no packet is bulk-class, and
introduce delayed ACK in the udp2raw shim so the ACK fraction approaches
0.33 rather than 0.5. We did not implement this defence in the present
work; it would require a patch to udp2raw and is logged as future work.
We expect that even with the defence in place, additional features (e.g.\
the precise distribution of small-packet sizes) would continue to leak.

\section{Related work}
\label{sec:related}

The literature on VPN-traffic and pluggable-transport detection is large
and actively contributed to. We highlight the work that is closest in
methodology or target.

\textbf{OpenVPN and Shadowsocks active probing.}
Xue~\textit{et al.}~\cite{xue2022openvpn} won Distinguished Paper at
USENIX~Security~'22 with a passive byte-and-length filter cascading into
an active prober that broke 34 of 41 commercial ``obfuscated'' OpenVPN
configurations. Alice~\textit{et~al.}~\cite{alice2020shadowsocks}
catalogued seven distinct active probe types used by the GFW against
Shadowsocks. Frolov~\textit{et~al.}~\cite{frolov2020probe} formalised
detection of probe-resistant proxies. Our C1 follows this lineage and
applies it to a previously unanalysed target.

\textbf{Recent protocol-agnostic methods.}
Xue~\textit{et~al.}~\cite{xue2024obfuscated} (USENIX~Security~'24) and
\cite{xue2025crosslayer} (NDSS~'25) propose obfuscated-tunnel detectors
that explicitly target ``any'' encapsulated proxy. The first identifies
nested TLS handshakes inside an outer encrypted stream; the second uses
RTT misalignment between transport and application layers. Both are out
of scope for our targets: WireGuard's outer encryption nullifies the
nested-TLS signal because no nested TLS exists, and our short, UDP-only
captures attenuate the cross-layer-RTT signal. Protocol-specific
structural features remain the available approach in this regime, and
that is our C2.

\textbf{Fully-encrypted-traffic detection by the GFW.}
Wu~\textit{et al.}~\cite{wu2023gfw} reverse-engineered the GFW's late-2021
entropy-and-popcount filter for ``fully encrypted'' protocols. Our positive
class is fully encrypted in the same sense, and the filter would likely
fire on bare WireGuard. Adopting their popcount feature for our setting
is straightforward and forms part of future work.

\textbf{ML on flow features for VPN identification.}
Mazel~\textit{et~al.}~\cite{mazel2022ml} (ANSSI) survey ML on flow metadata
for IPsec/OpenVPN/SSH/WireGuard tunnel identification, including bare
WireGuard but no obfuscated form. Recent ML work on
ISCXVPN2016~\cite{razooqi2025wavelet} continues to publish at ~98\%
binary accuracy with various feature engineering. We emphasise that our
two features are deliberately interpretable and protocol-grounded rather
than statistically engineered, and the resulting classifier is small
enough to read.

\textbf{WireGuard cryptography and formal verification.}
Dowling~\textit{et~al.}~\cite{dowling2018crypto} and a more recent
formal-methods analysis (USENIX Security~'25, in submission/published)
establish WireGuard's cryptographic correctness. Our work is orthogonal:
it concerns observability of the protocol's flow, not breakability of its
cryptography.

\textbf{Other obfuscation tools we did not study.}
AmneziaWG, swgp-go, wstunnel, and Cloak are alternative obfuscation
shims for WireGuard, each with its own design. We constrained our scope
to udp2raw because (a) it is by far the most-deployed of the set
(${\sim}$8.5k stars), (b) it is the only one with no academic analysis at
all, and (c) AmneziaWG~v2 in particular is documented to be a moving
target, with the v2 release in late 2025 introducing per-direction
ranged-header obfuscation that would invalidate any static feature
study published before its stabilisation. A comparative measurement
across the obfuscation ecosystem is logged as future work.

\section{Limitations and future work}
\label{sec:limits}

\textbf{Scope of evaluation.} All measurements are from a single
client/server pair in a controlled lab. We do not have an ISP-scale
deployment trace, and we explicitly do not claim our results transfer
unchanged to such an environment. The closely-related
Xue~\textit{et~al.}~\cite{xue2022openvpn} ISP deployment evaluation found
that some ``small'' lab signals attenuated in the field; we expect the
same to be possible here, and an ISP-scale evaluation is the
single most-impactful next step. We do not pursue it in the present work.

\textbf{Adversarial robustness.} Both contributions are static: a
sufficiently motivated udp2raw author can patch the upstream tool to
respond to P2/P4/P5 (closing C1) or to delay-ACK the faketcp segments
(closing C2). We have not engineered an arms race; we have characterised
the present unmodified deployment.

\textbf{Active-probing collateral risk.} An active prober that fires P3
(bogus-ACK) at non-target hosts is, in some legal regimes, a port
scan with mild hostility. In a real adversarial deployment the prober
would likely be run only against hosts already passively suspected.

\textbf{udp2raw versions.} Our analysis is of v20230206.0; the
multiplatform fork at version $\sim$2024 has minor differences in the
TCP option emulation and may answer P4 differently. We did not test it.

\textbf{Defences not yet evaluated.} A WireGuard MTU below the bulk
threshold combined with a delayed-ACK patch to udp2raw would substantially
weaken C2. We sketch the patch in \S\ref{sec:c2} but do not implement it.

\section{Conclusion}
\label{sec:conclusion}

The two-feature passive classifier and the five-probe active prober
together demonstrate that even when WireGuard's payload-byte fingerprint
is fully eliminated by udp2raw obfuscation, the wrapped tunnel remains
identifiable to a network observer with very modest tooling. Both
contributions are mechanical consequences of how the protocols are
implemented: WireGuard pads to MTU, udp2raw frames each datagram as a
faketcp segment that the receiver's kernel ACKs in 60 bytes, and udp2raw's
userspace TCP emulator silently drops packet shapes its parser does not
recognise.

For the practitioner deploying WireGuard for confidentiality on a trusted
network, this paper changes nothing. For the user behind a censorious
network who has installed udp2raw because the WireGuard byte-fingerprint
is being blocked, the paper documents that udp2raw alone is not sufficient
to render the tunnel passively or actively unobservable. We hope it is
useful as a baseline against which improved designs --- including the
delayed-ACK and MTU-cap defence we sketch above --- can be evaluated.

All scripts, datasets, configurations, and evidence files used in this
paper are available in the project repository.

\begin{thebibliography}{00}
\bibitem{donenfeld2017} J. A. Donenfeld, ``WireGuard: Next Generation
Kernel Network Tunnel,'' \textit{NDSS Symposium 2017}.

\bibitem{udp2raw} ``udp2raw-tunnel: A tunnel which turns UDP traffic into
encrypted UDP/FakeTCP/ICMP traffic,'' wangyu-, GitHub repository,
\url{https://github.com/wangyu-/udp2raw}, accessed 2026-05-09.

\bibitem{ghapi_udp2raw} GitHub REST API, release-asset download counts for
\texttt{wangyu-/udp2raw}, queried 2026-05-09 via
\url{https://api.github.com/repos/wangyu-/udp2raw/releases}.
Reproducible audit script in the project repository at
\texttt{notes/verify\_popularity.py}.

\bibitem{wgknown} ``Known Limitations,'' WireGuard project documentation,
\url{https://www.wireguard.com/known-limitations/}, accessed 2026-05-09.

\bibitem{amneziawg} ``AmneziaWG,'' Amnezia VPN project,
\url{https://github.com/amnezia-vpn/amneziawg-go}, accessed 2026-05-09.

\bibitem{xue2022openvpn} D. Xue, R. Ramesh, A. Jain, M. Kallitsis,
J. A. Halderman, J. R. Crandall, R. Ensafi, ``OpenVPN is Open to VPN
Fingerprinting,'' \textit{USENIX Security 2022}. (Distinguished Paper.)

\bibitem{xue2024obfuscated} D. Xue, M. Kallitsis, A. Houmansadr,
R. Ensafi, ``Fingerprinting Obfuscated Proxy Traffic with Encapsulated
TLS Handshakes,'' \textit{USENIX Security 2024}.

\bibitem{xue2025crosslayer} D. Xue, J. Stanley, P. Kumar, R. Ensafi,
``The Discriminative Power of Cross-layer RTTs in Fingerprinting Proxy
Traffic,'' \textit{NDSS 2025}.

\bibitem{wu2023gfw} M. Wu, J. Sippe, D. Sivakumar, J. Burg, P. Anderson,
X. Wang, K. Bock, A. Houmansadr, D. Levin, E. Wustrow, ``How the Great
Firewall of China Detects and Blocks Fully Encrypted Traffic,''
\textit{USENIX Security 2023}.

\bibitem{alice2020shadowsocks} ``Alice,'' ``Bob,'' ``Carol,'' J. Beznazwy,
A. Houmansadr, ``How China Detects and Blocks Shadowsocks,''
\textit{IMC 2020}.

\bibitem{frolov2020probe} S. Frolov, J. Wampler, E. Wustrow, ``Detecting
Probe-resistant Proxies,'' \textit{NDSS 2020}.

\bibitem{mazel2022ml} J. Mazel, M. Saudrais, A. Hervieu, ``ML-based tunnel
detection and tunneled application classification,'' arXiv:2201.10371,
2022.

\bibitem{razooqi2025wavelet} M. Razooqi, A. Pekar, ``Binary VPN Traffic
Detection Using Wavelet Features and Machine Learning,'' arXiv:2502.13804,
2025.

\bibitem{dowling2018crypto} B. Dowling, P. Rösler, J. Schwenk,
``A Cryptographic Analysis of the WireGuard Protocol,'' \textit{ACNS 2018}.

\bibitem{draper2016iscx} G. Draper-Gil, A. H. Lashkari, M. S. I. Mamun,
A. A. Ghorbani, ``Characterization of Encrypted and VPN Traffic using
Time-related Features,'' \textit{ICISSP 2016}. Dataset: ISCXVPN2016.

\bibitem{tegant_amnezia_2024} ``Iran's escalation of VPN-blocking
campaigns,'' Tegant blog and Amnezia issue tracker, 2024 (gray-literature
operational reports cited as evidence of deployment, not as primary
academic source).

\bibitem{scapy} P. Biondi and the Scapy community, ``Scapy: Packet
crafting for Python,'' \url{https://scapy.net}.
\end{thebibliography}

\end{document}
